---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# vida: The R package of the VIDA-DataBrew collaboration

## Installation


```r
if(!require(devtools)) install.packages("devtools")
install_github('databrew/vida')
```


```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.width = 4,
                      fig.height = 4)
library(dplyr)
library(tidyr)
library(ggplot2)
library(cism)

# knitr::knit_hooks$set(
#    error = function(x, options) {
#      paste('\n\n<div class="alert alert-danger">',
#            gsub('##', '\n', gsub('^##\ Error', '**Error**', x)),
#            '</div>', sep = '\n')
#    },
#    warning = function(x, options) {
#      paste('\n\n<div class="alert alert-warning">',
#            gsub('##', '\n', gsub('^##\ Warning:', '**Warning**', x)),
#            '</div>', sep = '\n')
#    },
#    message = function(x, options) {
#      paste('\n\n<div class="alert alert-info">',
#            gsub('##', '\n', x),
#            '</div>', sep = '\n')
#    }
# )
```

## About

`vida`, the R package of the VIDA-DataBrew collaboration, is a set of tools for transforming and combining verbal autopsy data from multiple sites and time periods.

## Objectives

The objectives of the R VIDA package are simple:

1. Make data "translation" straightforward and user friendly

2. Allow for translated data to be easily aggregated

3. Allow for aggregated data to be easily analyzed

## Set up


To use the `vida` package, you'll first need to install it. With a good internet connection, run the following from within R.

```r
if(!require(devtools)) install.packages("devtools")
install_github('databrew/vida')
```

Once you've installed the package, you can use its functionality in any R script by simply including the following line:

```r
library(vida)
```

## Basic use

### Format detection

The `vida` package can automatically detect the format of a dataset via the `detect_format` function. Here's an example of its use:

```{r}
library(vida)
# Load some fake data
fake <- vida::fake
# Detect the format
format <- detect_format(df = fake)
# Show the detected format
print(format)
```

### Data translation

The workhorse function of the package is `translate`, which translates data from any format to "babel" (the term we use to describe the standardized format for aggregating all datasets). Here's an example of its use:

```{r}
library(vida)
# Load some fake data
fake <- vida::fake
# Translate the data
translated <- translate(fake)
```

In addition to translating the data, the function prints information about how many variables were translatable/translated, as well as how many response categories were translatable/translated

## Developer details

### Raw data

Download the raw data into the "data_original" folder of this repo  [here](https://drive.google.com/drive/u/0/folders/1u-hRKraveojMvO9KAhYj_UzfYxkGn8nq).

### Kenya mapping information

Read about the Kenya files and mapping process [here](https://docs.google.com/document/d/13-DOozvzSteWy9JE_z5zmwbgAiJD2UFZrcjUbvpktZY/edit?usp=sharing).

### Mali mapping information
Download the mali_data folder into the main repository from [here](https://drive.google.com/drive/u/0/folders/1Nl6QrQhV3V_bYMWn7QK6jq96b-W3EJnp).
